---
title: "Bayesian Helper Functions"
author: "Dina Sinclair"
date: "May 15, 2018"
output: html_document
---
```{r, echo=FALSE }
# Set up workspace
rm(list = ls())
set.seed(1)

# Load packages
# require(data.table)
# library(ggplot2)
# require(Rcpp)
# require(RcppEigen) 
# library(gridExtra)
# library(dplyr)
# library(splitstackshape)
# library(sem)
# library(plyr)
# library(reshape)
# library(matrixStats)
# library(metafor)
# library(stringr)
```
## Basic Definitions

What we're working with here, drawing heavily from the HMCWG paper section 2.1:


* $Y_i$ is the point estimate in study $i$
* $\theta_i$ is the true effect in study $i$
* $\mu$ is the unknown mean of the distribution of $\theta_i$ 
* $\tau^2$ is the unknown variance of the distribution of $\theta_i$


$\mu$ and $\tau^2$ are known as *hyperparameters* since they model another one of our model variables, $\theta_i$.

$$\theta_i \sim N(\mu, \tau^2)$$



##First helper function: Suff Theta
```{r}
suff_theta <- function(Y, mu, tau2, sigma2){
  #vectorize for the whole sample size instead of one sample each run 
  n <- length(Y)
  m <- length(tau2)
  mean <- (matrix(Y, ncol = n, nrow = m, byrow = TRUE)*tau2 + 
             matrix(sigma2, ncol = n, nrow = m, byrow = TRUE)*mu) / 
    (matrix(sigma2, ncol = n, nrow = m, byrow = TRUE) + tau2)
  var <- matrix(sigma2, ncol = n, nrow = m, byrow = TRUE)*tau2 /
    (matrix(sigma2, ncol = n, nrow = m, byrow = TRUE) + tau2)
  #return the mean and variance for all e_i samples
  #in matrix with nrow = # of e_i and ncol = #samples
  return(list(mean = mean, var = var))
}
```

##Second helper function: Suff Mu
Here's the code:
```{r}
# sufficient statistics for P(mu|tau, Y)
suff_mu <- function(Y, sigma2, tau2){
  #one time each calculate, hard to vectorize
  lambda <- sum(Y/(sigma2 + tau2))
  omegai <- sum(1/(sigma2 + tau2))
  omega <- 1/omegai
  #return lambda and omega defined in the reference
  return(list(lambda = lambda, omega = omega))
}
```
Which mathematically is saying given $Y, \sigma^2$ and $\tau^2$,
\begin{align*}
\lambda &= \sum \frac{Y}{\sigma^2 + \tau^2} \\
\omega_i &= \sum \frac{1}{\sigma^2 + \tau^2} \\
\omega &= \frac{1}{\omega_i}
\end{align*}

Eva says the $\omega$ stuff is just to help with variance. Would that mean that 
$$V_\mu =\frac{1}{\sum \frac{1}{\sigma^2 + \tau^2}}=\omega$$
which matches up to equation (10) in section 2.2. Also from equation (10), we get that
$$\hat{\mu} = \frac{\lambda}{\omega_i} = \lambda \omega$$
so indeed, having $\lambda$ and $\omega$ is sufficient to derive $\mu$.

##Third helper function: Post Tau Squared
```{r}
# posterior samples from P(tau|Y), also get all the suff-stats for P(mu|tau, Y)
post_tau2 <- function(Y, sigma2, tau2, n.sims){
  # P(beta|tau, Y)
  N <- length(tau2) #grid length
  log_p <- rep(NA, N)
  lambda <- log_p
  omega <- log_p
  # run all the points we choose on the tau_grid
  for(ii in 1:N){
    suff <- suff_mu(Y, sigma2, tau2[ii])
    lambda[ii] <- suff$lambda*suff$omega
    omega[ii] <- suff$omega
    log_p[ii] <- 0.5*log(omega[ii]) - 0.5*sum(log(sigma2+tau2[ii])) - 
      0.5*sum((Y - lambda[ii])^2/(sigma2 + tau2[ii]))
  }
  log_p <- log_p - max(log_p) 
  p <- exp(log_p)
  p <- p/sum(p)
  index <- sample(1:N, n.sims, replace = T, prob = p)
  tau2 <- tau2[index]
  mean <- lambda[index]
  omega <- (sqrt(omega))[index]
  # return all the selected tau2 and mean (beta) and omega (for beta). 
  return(list(tau2 = tau2, mean = mean, omega = omega))
}
```

##Fourth helper function: Post Mix S
```{r}
#combine all the above functions for a joint sample of (theta, mu, tau)
#data is the raw data from the .csv selecting a certain group
#tau2 is in the grid
#Y <- data$treatmentcoefficient
post_mix_s <- function(data, Y, tau2, n_sim,onlymu = F){
  sigma2 <- (data$treatmentstandarderror)^2
  #first sample tau2
  tau2 <- post_tau2(Y, sigma2, tau2, n_sim)
  #also get mean and variance for beta
  mean <- tau2$mean
  omega <- tau2$omega
  tau2 <- tau2$tau2
  #calculate mu
  mu <- rnorm(n_sim)
  mu <- mu*omega + mean   
  if(onlymu){
    return(list(mu=mu))
  } else{
    #calculate suff-stats for theta
    theta <- suff_theta(Y, mu, tau2, sigma2)
    #sample theta
    theta <- array(rnorm(length(theta$mean), theta$mean, sqrt(theta$var)), 
                   dim = dim(theta$mean))
    #I2 calculation (reference page 161 "metafor.pdf")
    s2 <- (nrow(data)-1)*sum(1/sigma2)/((sum(1/sigma2))^2 - sum(1/sigma2^2))
    temp <- (mean(sqrt(tau2)))^2
    I2 <- 100*temp/(temp + s2)
    return(list(theta = t(theta), tau2 = tau2, mu = mu, I2 = I2))  
  }
  
}
```


