---
title: "Evaluating 2D REM"
author: "Dina Sinclair"
date: "December 12, 2018"
output: html_document
---

```{r}
library("rstan")

generateGroupedData <- function(I, mu, sd, N, SF=1){
  # Generates example dataset where data is grouped into N groups.
  # Note: Due to random group assignment some groups may have no data points assigned to them.
  # 
  # Arguments:
  #   I: the number of data points (cities)
  #   mu: the hyperparameter mean
  #   sd: the hyperparameter standard deviation
  #   N: the number of groups
  #   SF: the sigma factor, how wide spread we expect sigma to be

  # Generate mu/sd for each group
  mu_groups <- rnorm(N, mean = mu, sd = sd) # with theta ~ N(mu,sd)
  sd_groups <- SF * runif(N) # TODO(dsinc): correct sd generation to mirror stan code (lognormal?)
  G <- list(mu = mu_groups, sd=sd_groups)
  
  # Generate mu/sd for each study under a group
  group_assignment <- sample(1:N, I, replace=T)
  mu_studies <- numeric(I)
  sd_studies <- SF*runif(I)
  for (i in 1:I){
    mu_studies[i] <- rnorm(1,mean = G$mu[group_assignment[i]], sd = G$sd[group_assignment[i]])
  }
  Y <- list(mu = mu_studies, sd = sd_studies)
  
  # Save our generated input data together in a list
  generated_data <- list(I=I,
                              N=N,
                              Y_mean = Y$mu,
                              Y_sd = Y$sd,
                              groups=group_assignment)
  
  # Save our generated input data together in a list
  generated_data <- list(I = I,
                         N = N,
                         Y = Y,
                         G = G,
                         groups = group_assignment)
  
  # Return what we've generated
  return(generated_data)
}

parseGeneratedGroupedData <- function(generated_data){
  # Given a data input, outputs the needed format for the stan file
  # Save our generated input data together in a list
  stan_input_data_format <- list(I = generated_data$I,
                                 N = generated_data$N,
                                 Y_mean = generated_data$Y$mu,
                                 Y_sd = generated_data$Y$sd,
                                 groups=generated_data$groups)
  return(stan_input_data_format)
}
data <- parseGeneratedGroupedData(generateGroupedData(I=10,mu=0,sd=10,N=3))

data
```

```{r}
fit <- stan(file = 'randomEffectsModel2D.stan', 
            data = data,
            control = list(adapt_delta = 0.99,max_treedepth = 15),
            iter = 1000, chains = 2)
pairs(fit)
fit

```
I'm noticing that the mu/tau estimates aren't very good, even with 50 data points. Maybe try 500 or something --> even then, get my - -3.8 sd 50, tau=51 sd 90. Is that good or bad? Also, you see divergence in the data if you use one group, even if there are 50 data points in that group (don't see that divergence with 5 data points split among 3 groups, for example). But this does make things harder to test. Yeah why do I get so many divergent transitions when I make up numbers, ex below?
```{r}
data <- list(I=5, N=2, Y_mean=c(0.1,0.001,0.011,0,0.01), Y_sd=c(.01,.01,.01,.01,.01), groups=c(1,2,1,1,2))
fit <- stan(file = 'randomEffectsModel2D.stan', 
            data = data,
            control = list(adapt_delta = 0.99,max_treedepth = 15),
            iter = 1000, chains = 2)
fit
```


