---
title: "Evaluating 2D REM"
author: "Dina Sinclair"
date: "December 12, 2018"
output: html_document
---

```{r}
library("rstan")

generate_grouped_basic <- function(I,mu,tau,N,SF=1){
  # ARGUMENTS
  # I is the number of experiments (data points)
  # mu/tau are the overall mean/SD
  # N is the number of groups
  # SF is the sigma factor, how wide spread we expect sigma to be
  
  # Generate mu/tau for each group
  mu_groups <- rnorm(N, mean = mu, sd = tau) # with theta ~ N(mu,tau)
  tau_groups <- SF*runif(N) # with sigmaSq ~ U(0,1) #TODO I should switch this to a normal with var~SF, right??
  G <- list(mu = mu_groups, tau=tau_groups)
  
  # Generate mu/tau for each study under a group
  group_assignment <- sample(1:N, I, replace=T)
  mu_studies <- numeric(I)
  tau_studies <- SF*runif(I)
  for (i in 1:I){
    mu_studies[i] <- rnorm(1,mean = G$mu[group_assignment[i]], sd = G$tau[group_assignment[i]])
  }
  Y <- list(mu = mu_studies, tau = tau_studies)
  
  # Save our generated input data together in a list
  basic_dat_generated <- list(I=I,
                              N=N,
                              Y_mean = Y$mu,
                              Y_sd = Y$tau,
                              groups=group_assignment)
  
  # Display what we've generated
  return(basic_dat_generated)
}
data <- generate_grouped_basic(I=500,mu=0,tau=10,N=3)
#data
```

```{r}
fit <- stan(file = 'randomEffectsModel2D-new.stan', 
            data = data,
            control = list(adapt_delta = 0.99,max_treedepth = 15),
            iter = 1000, chains = 2)
pairs(fit)
fit

```
I'm noticing that the mu/tau estimates aren't very good, even with 50 data points. Maybe try 500 or something --> even then, get my - -3.8 sd 50, tau=51 sd 90. Is that good or bad? Also, you see divergence in the data if you use one group, even if there are 50 data points in that group (don't see that divergence with 5 data points split among 3 groups, for example). But this does make things harder to test. Yeah why do I get so many divergent transitions when I make up numbers, ex below?
```{r}
data <- list(I=5, N=2, Y_mean=c(0.1,0.001,0.011,0,0.01), Y_sd=c(.01,.01,.01,.01,.01), groups=c(1,2,1,1,2))
fit <- stan(file = 'randomEffectsModel2D-new.stan', 
            data = data,
            control = list(adapt_delta = 0.99,max_treedepth = 15),
            iter = 1000, chains = 2)
fit
```


