---
title: "BasicBayesian"
author: "Dina Sinclair"
date: "May 23, 2018"
output: html_document
---

```{r, warning = FALSE}
library("rstan")
```

The values I have inputted are currently completely arbitrary, with

 * $J$ is the number of individuals per study (**Note: this is currently set to be constant across all studies, which I need to change at some point because this is not realistic**)
 * $I$ is the number of studies
 * $Y_{i,j}$ is the data points from study $i$, individual $j$
 * $\sigma^2_{i,j}$ is the sampling variance of study $i$, individual $j$ 
 
We could choose values completely arbitrarily...
```{r}
basic_dat <- list(J = 3,
                    I =3,
                    Y = list(c(1,4,2),c(4,2,5),c(.1,.4,.4)),
                    sigmaSq = list(c(.3,.2,.6),c(1,2,3),c(.3,.3,.1))
)
```

Instead, it would be better to backtrack from mu and tau
Here we pick mu and tauSq to generate theta.

**Question: How can we pick good pretend sigmaSq values?**
```{r}
set.seed(17)
mu <- 10
tauSq <- 2
I <- 3
J <- 3
sigmaSq <- c(0.1,1,.2)

theta <- rnorm(I, mu, tauSq)
# This currently gives 1st row equal to theta(1), second row theta(2)... last row theta(I)
# Currently can have sigma vary by i (study) but not j (individual)
Y <- matrix( rnorm(I*J,mean=theta,sd=sigmaSq), I, J) 

basic_dat_generated <- list(J,I,Y,sigmaSq = list(sigmaSq, sigmaSq, sigmaSq))

theta
Y
```


Now we can run the model
```{r}
fit <- stan(file = 'basicBayesian.stan', data = basic_dat_generated, 
            iter = 10, chains = 4)
```

We get relatively reasonable backed out mu, tau, theta. Yay!
```{r}
fit
```



